---
title: "About Data"
author: "H24076150"
date: "2020/5/4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Remove NA
### If NA, ...
* is.na(X): easier for a vector.

* complete.cases(X): X could be a sequence of vectors, matrices and data frames.

* na.omit(X):

Examples:
```{r}
airquality[1:8, ]
dim(airquality)

no.na <- complete.cases(airquality)
head(airquality[no.na, ])

data.no.na.1 <- airquality[no.na, ]
dim(data.no.na.1)

data.no.na.2 <- na.omit(airquality)
dim(data.no.na.2)
```

## Data Processing
### Multiples datasets
* Using **cbind(...)** and **rbind(...)** to combine datasets.

* Using **cut(...)** to group or to label items.

For example, would like to label 5 groups by the variable "Wind".
```{r}
grouped1 <- cut(airquality[, 3], 5)
head(grouped1)

grouped2 <- cut(airquality[, 3], 5, labels = c("G1", "G2", "G3", "G4", "G5"))
head(grouped2)
```
### Summary the grouped data
(1) Summarize the number of each group.
```{r}
airquality$group <- grouped2
table(grouped2)
```
### Ordering the dataset
(2) After labeling the group, order the dataset by the group.

**order(...)**
```{r}
order.data.1 <- airquality[order(airquality$group), ]
head(order.data.1)
```
If there is no column for further sorting, it returns the result sorted by the index of the row.

(3) Keep ordering the dataset by the temprature.

**order(first, second, ...)**
```{}
order.data.2 <- airquality[order(airquality$group, airquality$Temp), ]
head(order.data.2)
```

## Package: reshape2
### Change the format from original one
Package: "reshape2"

For example, the dataset is as follows:
```{r}
game <- c("G1", "G2", "G3", "G4", "G5")
site <- c("M", "O", "L", "O", "L")
Lin <- c(15, 6, 26, 22, 18)
Jordan <- c(18, 32, 21, 25, 12)
Peter <- c(10, 6, 22, 9, 12)
ballgames <- data.frame(game, site, Lin, Jordan, Peter)
```

```{r, eval=FALSE}
install.packages("reshape2")
```

### melt(...)
USe **melt(data, id.vars)** to reshape the dataset.
```{r, warning=FALSE}
library(reshape2)
after.reshape <- melt(ballgames, id.vars = c("game", "site"))
after.reshape
name <- c("Lin", "Jordan", "Peter")
```

### dcast(...)
Use **dcast(dataset, formula, fun)** to show the specific data format.
```{r, echo=TRUE}
dcast(after.reshape, game + site ~ name, sum)

dcast(after.reshape, name ~ game, sum)

dcast(after.reshape, name ~ site, sum)

dcast(after.reshape, name + site ~ game, sum)
```

### Merge data
Two datasets:
```{r}
Lin.game <- ballgames[ballgames[, "Lin"] > 20, c("game", "site", "Lin")]
Jordan.game <- ballgames[ballgames[, "Jordan"] > 20, c("game", "site", "Jordan")]
Lin.game
Jordan.game
```

Using **merge(...)** to combine datasets to a new dataset by

(1) intersection

(2) union

By intersection:
```{r, echo=TRUE}
merge(Lin.game, Jordan.game)
```

By union:

set **all = TRUE** in merge(...)
```{r}
merge(Lin.game, Jordan.game, all = TRUE)
```

Merge data by the infromation of the specific dataset.

* merge(x, y, all.x = TRUE)

* merge(x, y, all.y = TRUE)
```{r}
merge(Lin.game, Jordan.game, all.x = TRUE)
merge(Lin.game, Jordan.game, all.y = TRUE)
```

## Regression
### linear model: lm(...)
* Regression is used to show the relationship between independent variables and dependent variables.

* Before data analysis, we should check the relation plot.
```{r}
pairs(airquality[, c(1, 3, 4)])
```

lm(formula, data)

formula: $y$ ~ $x_1+x_2$ (with intercept)

formula: $y$ ~ $x_1+x_2$ (no intercept)
```{r}
data1 <- airquality[1:120, ]
data2 <- airquality[121:153, ]
lm1 <- lm(Ozone ~ Temp, data = data1)
lm2 <- lm(Ozone ~ Temp - 1, data = data1)
```

### lm(...)
Result 1:
```{r}
summary(lm1)
```

Result 2:
```{r}
summary(lm2)
```

### Prediction by the fitted linear model
```{r}
y.predict <- predict(lm1, data2)
y.predict
```

### Performance of fitted model
```{r}
plot(airquality[, 4], airquality[, 1], pch = 19)
abline(lm1)
points(airquality[121:153, 4], y.predict, col = 2, pch = 19)
```

### Performance of prediction
```{r}
plot(airquality[121:153, 1], y.predict, pch = 19)
abline(0, 1)
```

## Package: dplyr
### %>% in package: dplyr
%>% means that apply functions to an object.
```{r, eval=FALSE}
install.packages("dplyr")
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
```

```{r}
y <- sample(1:100, 10)
mean(y)
y %>% mean()
min(y)
y %>% min()
```

### Multiple functions
```{r}
sample(1:100, 5) %>% mean()
sample(1:100, 10) %>% min()
```

### %>%
Write a code to report the last name of each member.
```{r}
nba_stars <- c("Stephen Curry", "Lebron James")
names <- strsplit(nba_stars, split = " ")
toupper(names[[1]][2])
```

```{r}
nba_stars %>%
  strsplit(, split = " ") %>%
  `[[` (1) %>%
  `[` (2) %>%
  toupper()
```