---
title: "Apply family"
author: "H24076150"
date: "2020/5/4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function: apply(...)
### Function: apply(...)
**apply(X, MARGIN, FUN, ...)** Returns a vector or array or list of values obtained by applying a function to margins of **an array or a matrix**.

* X: array or matrix (or a data frame)

* MARGIN: for a matrix 1 indicates rows, 2 indicates columns

* FUN: name of a function
```{r, echo=TRUE}
A <- matrix(rnorm(1500), 500, 3)
```
Sum of each row
```{r, results="hide"}
apply(A, 1, sum); rowSums(A)
```
Sum of each column
```{r, results="hide"}
apply(A, 2, sum); colSums(A)
```
Q: Write a function for the sum of each row.

### My own function
```{r, echo=TRUE}
row.col.sum <- function(X, margin = 1){
  size <- dim(X)
  if(margin == 1){
    res <- c()
    for(j in 1:size[1]){
      res <- c(res, sum(X[j, ]))
    }
  }else if(margin == 2){
    res <- c()
    for(j in 1:size[2]){
      res <- c(res, sum(X[, j]))
    }
  }else{
    stop("margin should be either 1 or 2.")
  }
  return(res)
}
```
Q: What is the meaning of "stop("margin should be either 1 or 2.")"?

### Check time-efficiency
Compare the evaluation time of multiple functions.
```{r, echo=TRUE}
library(rbenchmark)
benchmark(replications = rep(1000, 1), apply(A, 1, sum), rowSums(A), row.col.sum(A, 1), columns = c("test", "elapsed", "relative"))
```

### What if there are NA's?
```{r, echo=TRUE}
B <- matrix(sample(1:100, 15), 5, 3)
B[1, 2] <- NA
B[3, 1] <- NA
B <- data.frame(B)
colnames(B) <- c("factor 1", "factor 2", "factor 3")
apply(B, 2, mean)
```
Recall: mean(..., na.rm = TRUE)
```{r, echo=TRUE}
apply(B, 2, mean, na.rm = TRUE)
```

### Try other function in apply(X, MARGIN, FUN,)
```{r, echo=TRUE}
apply(B, 2, max)
apply(B, 2, min, na.rm = TRUE)
```

### NOTE: improving the code above
```{r, echo=TRUE}
row.col.sum.ver2 <- function(X, margin = 1){
  size <- dim(X)
  if(margin == 1){
    res <- rep(NA, size[1])
    for(j in 1:size[1]){
      res[j] <- sum(X[j, ])
    }
  }else if(margin == 2){
    res <- rep(NA, size[2])
    for(j in 1:size[2]){
      res[j] <- sum(X[, j])
    }
  }else{
    stop("margin should be either 1 or 2.")
  }
  return(res)
}
```
Note: Change the initial vector!

### Compare the efficient
```{r, echo=TRUE}
benchmark(replications = rep(1000, 1), row.col.sum(A, 1), row.col.sum.ver2(A, 1), rowSums(A), columns = c("test", "elapsed", "relative"))
```

## Function: lapply(...)
### Function: lapply(...)
**lapply(X, FUN, ...)**

Returns a list of the same length as X, each element of which is the result of applying FUN to the corresponding element of X.

* X: vector or data frame or list

* FUN: name of a function
```{r, echo=TRUE, error=TRUE}
C <- list(a = 1:10, beta = exp(-3:3))
apply(C, FUN = mean)

lapply(C, mean)

par(mfrow = c(1, 2))
lapply(C, hist)
```

## Function: sapply(...)
### Function: sapply(...)
**sapply(X, FUN, ..., simplify = TRUE**

Similarlly to lapply(X, FUN, ...) if sapply(x, f, simplify = FALSE, USE.NAMES = FALSE)

* X: vector or data frame or list

* FUN: name of a function

* simplify: logical; TRUE will return an array or a matrix with the same length of oytput
```{r, echo=TRUE}
sapply(C, summary)
sapply(C, summary, simplify = FALSE)
lapply(C, summary)
```

### If there are different lengths in the result?
Using **apply**, it returns a list.
```{r, echo=TRUE}
D <- list(a = sample(1:10, 10 ,replace = TRUE), b = sample(1:5, 20 ,replace = TRUE))
sapply(D, unique)
sapply(D, table)
```

## Others
tapply(...)