---
title: "Control Flow"
author: "H24076150"
date: "2020/4/30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## The for() loop
### For() loop
When to use?

* See patterns of behaviour, not just a single instance.
* The for() statement allows one to specify that a certain operation should be repeated a fixed number of times.

**Syntax**
<center>for (index in vector) commands,</center>

where

* index: the index is changed each time
* vector: it includes all possible cases
* commands: actions

### Example of for() [Version 1]
Construct a vector as
<center>$X_{n+1}=2X_n+3,$</center>

where $X_0=3,n=1,2,3,4,5,6,7.$
```{r, echo=TRUE}
vec.X.v1 <- c()
Xn <- 3
for(n in 1:7){
  Xn <- 2*Xn + 3
  vec.X.v1 <- c(vec.X.v1, Xn)
}
vec.X.v1
```
### Example of for() [Version 2]
```{r, echo=TRUE}
vec.X.v2 <- rep(NA, 8)
vec.X.v2[1] <- 3
for(n in 1:7){
  Xn <- 2*Xn + 3
  vec.X.v2[n+1] <- 2 * vec.X.v2[n] + 3
}
vec.X.v2
vec.X.v2[2:8]
vec.X.v2[-1]
```
### Evaluate n!
<center>$n!=n*(n-1)*(n-2)*...*2*1.$</center>

Try $n=10$.
```{r, echo=TRUE}
n <- 10
ans <- 1
for(j in n:1){
  ans <- ans * j
}
ans
factorial(n)
```
### Random number generator
Random samples from [0,1] by the linear congruntial generator.

<center>$x_n=(ax_{n-1}+c)$ mod $m$.</center>

Let $x_0=3,a=5,c=7$, and $m=200$. Find $x_i$, where $i=1,...,n$. What are the values after all in $X_n$ devided by $m$?
```{r, echo=TRUE}
x0 <- 3; a <- 5; c <- 7; m <- 501
X <- rep(NA,10)
for(j in 1:10){
  x0 <- (a*x0+c) %% m
  X[j] <- x0
}
X
round(X/m, 2)

x0 <- 3; a <- 5; c <- 7; m <- 501
X <- rep(NA,11)
X[1] <- x0
for(j in 1:10){
 X[j + 1] <- (a*X[j] + c) %% m
}
X[2:11]
round(X/m, 2)
```
### Try the Fibonacci sequence
Find the first 12 values in the Fibonacci sequence.

Let $F_1=1$ and $F_2=1$.

<center>$F_n=F_{n-1}+F_{n-2},$ $n=3,4,5,...,12$</center>

or

<center>$F_{n+2}=F_{n+1}+F_n,$ $n=1,2,3,...,10$</center>
```{r, echo=TRUE}
f1 <- 1 ; f2 <- 1
Fibonacci <- c(f1)
while(f2<100){
  Fibonacci <- c(Fibonacci,f2)
  f1 <- Fibonacci[length(Fibonacci)-1]
  f2 <- f1 + Fibonacci[length(Fibonacci)]
}
Fibonacci
```
### For loop in matrix
Given a matrix

<center>$$X=
\begin{bmatrix}
3 & 2 & 4 & 5 & 3\\
2 & 6 & 8 & 10 & 4\\
1 & 4 & 6 & 8 & 3\\
2 & 1 & 2 & 6 & 4
\end{bmatrix}
$$</center>

(1) Evaluate the mean for each row.
```{r, echo=TRUE}
A <- matrix(c(3, 2, 1, 2, 2, 6, 4, 1, 4, 8, 6, 2, 5, 10, 8, 6, 3, 4, 3, 4), 4, 5)
nr <- dim(A)[1]
mr <- c()
for(j in 1:nr){
  mr <- c(mr, mean(A[j, ]))
}
mr
```
(2) Evaluate the mean for each column.
```{r, echo=TRUE}
A <- matrix(c(3, 2, 1, 2, 2, 6, 4, 1, 4, 8, 6, 2, 5, 10, 8, 6, 3, 4, 3, 4), 4, 5)
nc <- dim(A)[2]
mc <- c()
for(j in 1:nc){
  mc <- c(mc, mean(A[, j]))
}
mc
```
Given a matrix

<center>$$X=
\begin{bmatrix}
3 & 2 & 4 & 5 & 3\\
2 & 6 & 8 & 10 & 4\\
1 & 4 & 6 & 8 & 3\\
2 & 1 & 2 & 6 & 4\\
4 & 0 & 3 & 2 & 1
\end{bmatrix}
$$</center>

extract the diagonal of the matrix.
```{r, echo=TRUE}
B <- matrix(c(3, 2, 1, 2, 4, 2, 6, 4, 1, 0, 4, 8, 6, 2, 3, 5, 10, 8, 6, 2, 3, 4, 3, 4, 1), 5, 5)
n <- dim(B)[1]
d <- c()
for(j in 1:n){
  d <- c(d, B[j, j])
}
d
```
## The if() statement
### if()
**Syntax**

if (condition) {commands when TRUE}

if (condition) {commands when TRUE} else {commands when FALSE}

* This statement causes a set of commands to be invoked if condition evaluates to TRUE.

* The else part is **optional**, and provides an alternative set of commands which are to be invoked in case the logical variable is FALSE.


The highly recommended to type the format as follow:

* One condition:

  if (condition) {

  commands when TRUE

  } else {

  commands when FALSE

  }

* Multiple conditions:

  if (condition1) {

  commands when condition1 is TRUE

  } else if (condition2) {

  commands when condition2 is TRUE

  } else {

  commands when condition1 and condition2 are FALSE

  }

### Try to find the locations of NA
Given $C=${3,5,NA,12,0.2,NA,-2}, find the locations of NA.
```{r, echo=TRUE}
C <- c(3, 4, NA, 12, 0.2, NA, -2)
for(i in 1:7){
  if(is.na(C[i])) print(i)
}
```
Extract the locations of NA to a new vector
```{r, echo=TRUE}
na.loc <- c()
for(i in 1:7){
  if(is.na(C[i])) na.loc <- c(na.loc, i)
}
na.loc
```
### Count the number of each class
basket$=${"r","b","y","y","y","r","r","b"}
```{r, echo=TRUE}
basket <- c("r","b","y","y","y","r","r","b")
n.r <- 0; n.b <- 0; n.y <- 0
for(i in 1:length(basket)){
  if(basket[i] == "r"){
    n.r <- n.r + 1
  }else if(basket[i] == "b"){
    n.b <- n.b + 1
  }else{
    n.y <- n.y + 1
  }
}
c(n.r, n.b, n.y)
```
### Count the number of each class by vectorization
```{r, echo=TRUE}
ind.r <- basket == "r"
ind.b <- basket == "b"
ind.y <- basket == "y"
c(sum(ind.r), sum(ind.b), sum(ind.y))
```
### Re-assign the value to the class
Let "r" be 1, "b" be 2, and "y" be 3.
```{r, echo=TRUE}
v.basket <- rep(NA, length(basket))
for(i in 1:length(basket)){
  if(basket[i] == "r"){
    v.basket[i] <- 1
  }else if(basket[i] == "b"){
    v.basket[i] <- 2
  }else{
    v.basket[i] <- 3
  }
}
v.basket
```
### Re-assign the value to the class by vectorization
Let "r" be 1, "b" be 2, and "y" be 3.
```{r, echo=TRUE}
v.basket <- rep(NA, length(basket))
ind.r <- basket == "r"
ind.b <- basket == "b"
ind.y <- basket == "y"
v.basket[ind.r] <- 1
v.basket[ind.b] <- 2
v.basket[ind.y] <- 3
v.basket
```
### Re-assign the value to the class by ifelse()
ifelse(condition, commands when TRUE, commands when FALSE)

Let "r" be 1, "b" be 2, and "y" be 3.
```{r, echo=TRUE}
v.basket <- ifelse(basket == "r", 1, ifelse(basket == "b", 2, 3))
v.basket
```
## The while() loop
### while()
The pattern of repetition isn't known in advance and we need to do some calculations and keep going as long as a condition holds. The while() statement accomplishes this.

**Syntax**

<center>while (condition) statements</center>

If the condition evaluates to FALSE, nothing more is done. If it evaluates to TRUE the statements are executef, condition is evaluated again, and the process is repeated.

### Relationship between for() and while()
* for
```{r, echo=TRUE}
for(j in 1:8){
  print(j)
}
```
* while(): Need the initial, condition of stopping, and iteration or replacement
```{r, echo=TRUE}
j <- 1
while(j <= 8){
  print(j)
  j <- j + 1
}
```
### Fibonacci number (1)
List all Fibonacci numbers less than 100.

Can you know the location that a Fibonacci number is small than 100 in advance?

Use for() or while()?
```{r, echo=TRUE}
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
while (Fib2 < 100){
  Fibonacci <- c(Fibonacci, Fib2)
  Fib1 <- Fibonacci[length(Fibonacci) - 1]
  Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
}
Fibonacci
```
### Flip a coin
Let 0 be a tail and 1 be a head. Count the total number of flipping until each success occurs.
```{r, echo=TRUE}
coin <- c(0, 0, 0, 0, 0, 0, 1, 0, 1, 0, 1, 0, 1, 1, 0, 1, 1, 0, 0, 0, 1, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1)
N <- c(); j <- 1
while(j <= length(coin)){
  n.fail <- 0
  while(coin[j] != 1){
    n.fail <- n.fail + 1
    j <- j + 1
  }
  N <- c(N, n.fail + 1)
  j <- j + 1
}
N
```
## break, repeat, and next
### break & repeat
* if (condition) break

The break statement causes the loop to terminate immediately. Break statements can also be used in for() and while() loops.

* repeat

Use "repeat" to repeat until we execute a break statement.

repeat {statements}

This causes the statements to be repeated endlessly.

* next

The next statement causes control to return immediately to the top of the loop; it can also be used in any loop.
### Fibonacci number (2)
List all Fibonacci numbers less than 100. Use for() and if(condition) break.
```{r, echo=TRUE}
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
for(j in 1:1000){
  if(Fib2 < 100){  
    Fibonacci <- c(Fibonacci, Fib2)
    Fib1 <- Fibonacci[length(Fibonacci) - 1]
    Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
  }else{
    break
  }
}
Fibonacci
```
### Fibonacci number (3)
List all Fibonacci numbers less than 100. Use repeat{} and break.
```{r, echo=TRUE}
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
repeat
{
  if(Fib2 < 100){  
    Fibonacci <- c(Fibonacci, Fib2)
    Fib1 <- Fibonacci[length(Fibonacci) - 1]
    Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
  }else{
    break
  }
}
Fibonacci
```
### Fibonacci number (4)
List all Fibonacci numbers less than 100. Use repeat{} and break.
```{r, echo=TRUE}
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
repeat
{
  if(Fib2 >= 100) break  
    Fibonacci <- c(Fibonacci, Fib2)
    Fib1 <- Fibonacci[length(Fibonacci) - 1]
    Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
}
Fibonacci
```
### Fibonacci number (5)
List all Fibonacci numbers less than 100. Use repeat{} and next.
```{r, echo=TRUE}
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
for(j in 1:1000){
  if(Fib2 >= 100) next
    Fibonacci <- c(Fibonacci, Fib2)
    Fib1 <- Fibonacci[length(Fibonacci) - 1]
    Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
}
Fibonacci
```