---
title: "Functions"
author: "H24076150"
date: "2020/5/2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Use source()
### source()
(1) Creat a new script for your function or codes.
(2) Save the script named as "XXX.R" and use scource() to run the code.
(3) Remember to use print() or cat() so that your final answer will be shown.

Example:
```{r, echo=TRUE}
x <- c(0.8933, 0.2344, 0.0764, 0.7736)
x.percent <- round(100*x, digits = 1)
x.show <- paste(x.percent, "%", sep = "")
print(x.show)
cat(x.show)
```
### print() & cat()
**print()** shows with " " between characters.

**cat()** shows without " " between characters. You can add "\\n" to make characters in a new line.
```{r, echo=TRUE}
x <- c("A", "B", "C", "D")
print(x)
cat(x)
cat(x[1], "\n", x[2], "\n", sep = "")
```
### Another example by source()
Find all perfect numbers smaller than 100. Show the results by **print()**.
```{r, echo=TRUE}
k <- 1
while(k < 100){
  fac <- c()
  for(j in 1:k){
    if((k %% j) == 0) fac <- c(fac, j)
  }
  if(k == sum(fac[-length(fac)])){
    out <- paste(paste(k, " = ", sep = ""),
                 paste(fac[-length(fac)], collapse = " + "), sep = " ")
    print(out)
  }
  k <- k + 1
}
```
Find all perfect numbers smaller than 100. Show the results by **cat()**.
```{r, echo=TRUE}
k <- 1
while(k < 100){
  fac <- c()
  for(j in 1:k){
    if((k %% j) == 0) fac <- c(fac, j)
  }
  if(k == sum(fac[-length(fac)])){
    cat(paste(k, " = ", sep = ""),
        paste(fac[-length(fac)], collapse = " + "), "\n")
  }
  k <- k + 1
}
```
Find all perfect numbers smaller than 100. Show the results by **cat()** but without "\\n".
```{r, echo=TRUE}
k <- 1
while(k < 100){
  fac <- c()
  for(j in 1:k){
    if((k %% j) == 0) fac <- c(fac, j)
  }
  if(k == sum(fac[-length(fac)])){
    cat(paste(k, " = ", sep = ""),
        paste(fac[-length(fac)], collapse = " + "))
  }
  k <- k + 1
}
```
## Use function() with one single input
### function()
Use **function()** when you only have one main code for evaluation and multiple possible input, like

<center>$f(x)=3x+2, x=2,3,4$.</center>
$f$ means a name of a function of input $x$ when $x=2,3,4$.
```{r, echo=TRUE}
f <- function(x){
  3*x + 2
}
f(2); f(3); f(4)
```
### Example for function()
```{r, echo=TRUE}
make.percent <- function(x){
  x.percent <- round(100*x, digits = 1)
  x.show <- paste(x.percent, "%", sep = "")
  cat(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans <- make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans ## Why does ans show NULL?
```
### return() in the function()
You must tell what will be return when you use the function by **return()** or **invisible()**. If you don't put **return()** or at the end of the function, it will treat the last line as the result automatically.
```{r, echo=TRUE}
f <- function(x){
  return(3*x + 2)
}
f(2)
f <- function(x){
  res <- (3*x + 2)
}
f(2) ## why return nothing?
a <- f(2); a
```
### Example: percent by function() and return()
```{r, echo=TRUE}
make.percent <- function(x){
  x.percent <- round(100*x, digits = 1)
  x.show <- paste(x.percent, "%", sep = "")
  return(x.show) ## compare to cat(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans <- make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans
```
### Example: percent by function() and invisible()
```{r, echo=TRUE}
make.percent <- function(x){
  x.percent <- round(100*x, digits = 1)
  x.show <- paste(x.percent, "%", sep = "")
  invisible(x.show) ## compare to return(x.show)
}

make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans <- make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans
```
### Perfect number
Find all perfect numbers smaller than 10000.Show the results by **cat()**, named **perect.number**. After getting all the perfect numbers, count the number of all perfect numbers smaller than 10000.
```{r, echo=TRUE}
perfect.number <- function(n){
  k <- 1
  out <- c()
  while(k < n){
    fac <- c ()
    for(j in 1:k){
      if((k %% j) == 0) fac <- c(fac, j)
    }  
    if(sum(fac[-length(fac)])==k) out <- c(out, k)
    k <- k + 1
  }
  cat(out)
}
perfect.number(10000)
ans <- perfect.number(10000)
length(ans)
```
### Remark for the functions
* The name of a function is also a type of an object.

* If there are a vector and a function with the same name, then the type of the object will be the last one.

Example
```{r, echo=TRUE}
objA <- function(x){
  return(2*x)
}
objA ## give a function
objA <- objA(4)
objA ## give a value
```
## Use function() with multiple input variables
### function()
Use **function()**when yoy only have one main code for ebaluation and multiple possible input, like

<center>$f(x,y)=x^2+y^2,$ $x,y=2,3,4.$</center>

$f$ means a name of a function of input $x$ and $y$ when $x,y=2,3,4$.
```{r, echo=TRUE}
f <- function(x, y){
  x^2 + y^2
}
f(2, 2); f(2, 3); f(2, 4)
```
### Example: percent
```{r, echo=TRUE}
make.percent <- function(x, x.digits){
  x.percent <- round(100*x, digits = x.digits)
  x.show <- paste(x.percent, "%", sep = "")
  return(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736), 0)
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736), 1)
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736), 2)
```
### Example: perfect number
```{r, echo=TRUE}
perfect.number <- function(a, b){
  k <- a
  res <- c()
  while(k < b){
    fac <- c ()
    for(j in 1:k){
      if((k %% j) == 0) fac <- c(fac,j)
    }  
    if(sum(fac[-length(fac)]) == k) res <- c(res, k)
    k <- k + 1
  }
  return(res)
}
perfect.number(10, 10000)
```
### Set a default in the function
Add the settings og default in the arguments of input variables.

<center> f <- function(x = 1, y = 10){...}</center>

Example: percent
```{r, echo=TRUE}
make.percent <- function(x, x.digits = 0){
  x.percent <- round(100*x, digits = x.digits)
  x.show <- paste(x.percent, "%", sep = "")
  return(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736), 0)
```
## Use function() and return multiple output
### Introduction to list()
**list()** is a collection of multiple types of objects, like vectors, characters, numbers, lists,...

Example:
```{r, echo=TRUE}
X <- list(a = 3, b = c(1, 2, 3, 4, 5), c = c("A", "B"))
X
```
### Extract an element from a list
USe **$** or **[[]]**.

* **$**: the name of the element.
* **[[]]**: the location or the name of the element.
```{r, echo=TRUE}
X <- list(a = 3, b = c(1, 2, 3, 4, 5), c = c("A", "B"))
X$a
X[[2]]
X[["c"]]
```
### Creat a list
```{r, echo=TRUE}
try.list <- list() ## a null list
length(try.list)
length(try.list) <- 2  ## assign the size to the list
try.list[[1]] <- seq(0, 10, length.out = 3) ## without name
try.list$square <- seq(0, 10, length.out = 3)^2
try.list
```
### Extract element from the list
```{r, echo=TRUE}
try.list[[3]]
try.list$square
names(try.list)
length(try.list)
```
### Use function() and list()
Try to use **list()** to return more than one output.

Example: perfect number
```{r, echo=TRUE}
perfect.number <- function(a, b){
  k <- a
  res <- c()
  ans <- list()
  while(k < b){
    fac <- c ()
    for(j in 1:k){
      if((k %% j) == 0) fac <- c(fac,j)
    }  
    if(sum(fac[-length(fac)]) == k){
      res <- c(res, k)
      ans[[paste(k)]] <- fac[-length(fac)]
    }
      k <- k + 1
  }
  ans$number <- res
  return(ans)
}
```
### Output: perfect number
```{r, echo=TRUE}
Ans <- perfect.number(1, 100)
Ans
Ans[["28"]]
```
### Partially shown multiple output
Use **invisible()** to hide some output.

Example: perfect number
```{r, echo=TRUE}
perfect.number <- function(a, b){
  k <- a
  res <- c()
  ans <- list()
  while(k < b){
    fac <- c ()
    for(j in 1:k){
      if((k %% j) == 0) fac <- c(fac,j)
    }  
    if(sum(fac[-length(fac)]) == k){
      res <- c(res, k)
      ans[[paste(k)]] <- fac[-length(fac)]
    }
      k <- k + 1
  }
  print(res)
  ans$number <- res
  return(invisible(ans))
}
```
### Output: perfect number
```{r, echo=TRUE}
Ans <- perfect.number(1, 100)
Ans
length(Ans)
```