---
title: "Introduction"
author: "H24076150"
date: "2020/4/29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Calculator
```{r, echo=TRUE}
5 + 3
5 * 3
5 / 3
5 ^ 3
5 %% 3
5%/%3
```
## Sequence
```{r, echo=TRUE}
1:10
seq(1, 10, 2)
seq(1, 10, length.out = 6)
```
## Basic features in R
```{r, echo=TRUE}
a <- 3 ^ 5
b = 3 ^ 5
a
b
```
## Calculation by the set values
```{r, echo=TRUE}
a <- 1000
b <- 4
a * b - a
# If you don't assign a value to a name of variable, then you will get the error message.
```
## Special values in R 
```{r, echo=TRUE}
NA
NaN   ###0/0
Inf   ###2/0
-Inf  ###-2/0
pi
1e-3
```
### More example of special values in R
```{r, echo=TRUE}
Inf + 100
Inf - Inf + 10
NaN + Inf
NA + Inf
NaN + NA
```
## Name an object
* Avoid the following words:
  + break; else; FALSE; F; TRUE; T; for; function; if; Inf; NA; NaN; next; repeat; return; while
* Name a meaningful object by "." or "_"
  + score1, score2
  + score.John, score.Jack

## More on calculation
```{r, echo=TRUE}
# Absolute value
abs(-2)
# Exponential
exp(10)
# logarithm
log(2); log2(2); log(10); log10(10)
log(10, base = 10); log10(10)
log(10, base = exp(1)); log(10)
```
## Other functions
"^" or "**" means the degree of the power.
```{r, echo=TRUE}
2 ^ 3; 2 ** 3
2 ^ 0.5; sqrt(2)
```
Trigonometric function: sin(), cos(), tan(), asin(), acos(), atan()

## Approximated values (1)
```{r, echo=TRUE}
round(125.78253)
round(125.78253, digits = 3)
round(125.78253, digits = -2)
round(83524587, digits = -3); round(83524587, digits = -5)
```
## Approximated values (2) 
```{r, echo=TRUE}
signif(125.78253, digits = 2)
signif(83524587, digits = 3); signif(83524587, digits = 5)
floor(125.78253)
ceiling(125.78253)
trunc(125.78253)
```
## R is not just a calculator
There are a lot of functions you can use. Let's see some related to Statistics.
```{r, echo=TRUE}
x <- 1:25
mean(x)
var(x)
median(x)
quantile(x, c(0.1, 0.5, 0.7))
```