---
title: "Matrix"
author: "H24076150"
date: "2020/4/30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Construct a matrix
### Construct a matrix by row
A matrix is a generalization containung multiple vectors.
```{r, echo=TRUE}
a <- 1:5; b <- 6:10;
c <- 11:15; d <- 16:20
```
How to construct a matrix?
```{r, echo=TRUE}
M1 <- rbind(a, b, c, d)
M1
dim(M1)
```
### Construct a matrix by column
```{r, echo=TRUE}
M2 <- cbind(a, b, c, d)
M2
dim(M2)
```
### Construct a matrix by a long vector
A matrix is a generalization containung multiple vectors.
```{r, echo=TRUE}
X <- 1:20
matrix(X, nrow = 4, ncol = 5)
matrix(X, nrow = 4, ncol = 5, byrow = TRUE)
matrix(X, nrow = 5, ncol = 4)
matrix(X, nrow = 5, ncol = 4, byrow = TRUE)
```
### Features for a matrix 
```{r, echo=TRUE}
M1
dim(M1)
colnames(M1); rownames(M1)
colnames(M1) <- c("V1", "V2", "V3", "V4", "V5")
```
### Names in a matrix
```{r, echo=TRUE}
dimnames(M1)
dimnames(M2)
```
## Extract elements from a matrix
We still use "[]" to extract elements from a matrix, but using a "," to identify the locations of row and column sperately.
```{r, echo=TRUE}
M1[2, 4]; M1[3, 5]
M1[c(2, 4), 3]; M1[c(2, 4), c(3, 5)]
```
If you think these elements are from a vecotor but a matrix, you will type:
```{r, echo=TRUE}
M1
M1[5]
M1[13]
M2
M2[5]
M2[13]
```
### Extract a row or a column from a matrix
```{r, echo=TRUE}
M1[, 4]; M1[3, ]
M1[c(2, 4), ]; M1[, c(3, 5)]
```
### Extract elements by names
Recall M1. The names of rows are a, b, c, and d. Take the rows of b and c.
```{r, echo=TRUE}
M1[c("b", "c"),]
```
The names of columns are V1, V2, V3, V4, and V5. Take the rows of b and c and the columns of V2 and V5.
```{r, echo=TRUE}
M1[c("b", "c"),c("V2", "V5")]
```