---
title: "R Graphics"
author: "H24076150"
date: "2020/5/3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Statistical Figures
### Graphics for descripitive statistics
Given an arbitrary vector x:

(1) Histogram: hist(x)

(2) Boxplot: boxplot(x)

(3) Density plot: plot(density(x))
```{r, echo=TRUE}
x.norm <- rnorm(100, mean = 0, sd = 1)
hist(x.norm)
boxplot(x.norm)
plot(density(x.norm))
```

For discrete variables x:

(1) Barchart: barplot(x)

(2) Piechart: pie(x)
```{r, echo=TRUE}
x <- sample(c(1:8), 100, replace = TRUE, prob = c(8:1))
x <- table(x)
names(x) <- c("A", "B", "C", "D", "E", "F", "G", "H")
barplot(x)
pie(x)
```

Scatter plot is for the correlation between two variables.

Use **plot(x, y)**
```{r, echo=TRUE}
x <- rnorm(100)
y <- 3*x + rnorm(100, 0, 1)
plot(x, y)
cor(x, y)
```
### Scatter plots for multiple variables
Use **pairs(...)**
```{r, echo=TRUE}
X <- matrix(rnorm(30), 10, 3) ## use matrix
pairs(X, pch = 19)
```

### Boxplot for multiple categories
```{r, echo=TRUE}
X <- matrix(rnorm(50), 10, 5) ## use matrix
boxplot(X)
X <- data.frame(group = rep(c("A", "B", "C", "D", "E"), each = 10), res = rnorm(50)) ## use data.frame
boxplot(res ~ group, data = X)
```

## High-level plotting functions: Creat a plot region with specifical items
### General plot function: plot(...)
```{}
plot(x, y, 
     type = "p",               ## types of the plot
     pch = ,                   ## the symbol in the plot
     main = NULL,              ## the title of the plot
     lty = ,                   ## the style of a line
     lwd = ,                   ## the width of a line
     cex = ,                   ## the size of an text
     bty = "o",                ## the box around the plot
     col = ,                   ## the color of an object
     xlim = NULL, ylin = NULL, ## the range of an axis
     sub = NULL,               ## the sub-title of the plot
     xlab = NULL, ylab = NULL, ## the name of an axis
     cex.main =                ## the size of the title
     )
```
### plot(..., type = ...)
```{r, echo=TRUE}
type = c("p", "l", "b", "s", "o", "h", "n")
```
* "p": points
* "l": lines
* "b": both points and lines
* "o": lines pass through points
* "h": the lines between points and x-axis
* "n": no content in the plot but plot region
```{r, echo=TRUE}
par(mfrow = c(2, 3)) ## split plot ranges
plot(x, y, type = "p", main = "p")
plot(x, y, type = "l", main = "l")
plot(x, y, type = "b", main = "b")
plot(x, y, type = "o", main = "o")
plot(x, y, type = "h", main = "h")
plot(x, y, type = "n", main = "n")
```

### plot(..., pch = ...)
Choose a symbols for points.
```{}
pch = c(1, 2, ..., 25, "*", ".", "o", "0", "0", "+", "-", "|", "%", "#")
```
### plot(..., cex = ...)
Sizes in a plot: the number of times size of the default one

* Size of points: **cex = 1.5**

* Size of the title: **cex.main = 2**

* Size of the labels of ticks of an axis: **cex.axis = 2**

* Size of the labels of the name of an axis: **cex.lab = 2**

### plot(..., pch = ..., cex = ....)
```{r, echo=TRUE}
par(mfrow = c(2, 3)) ## split plot ranges
plot(x, y, pch = 1, main = "pch = 1")
plot(x, y, pch = 3, main = "pch = 3", cex.main = 2)
plot(x, y, pch = 10, main = "pch = 10", cex.axis = 2)
plot(x, y, pch = 15, main = "pch = 15", cex = 2, cex.main = 2)
plot(x, y, pch = 19, main = "pch = 19", cex = 2, cex.axis = 2)
plot(x, y, pch = "#", main = "pch = #", cex = 2, cex.axis = 2, cex.lab = 2, cex.main = 2)
```

### plot(..., type = "l", lty = ..., lwd = ...)
Lines in the plot:

* The style of lines: **lty = 1, 2, ..., 7**

* The width of lines: **lwd = 1, 1.25, ...**
```{r, echo=TRUE}
t <- 1:20
y <- rnorm(20)
par(mfrow = c(2, 3)) ## split plot ranges
plot(t, y, type = "l")
plot(t, y, type = "l", lty = 2)
plot(t, y, type = "l", lty = 2, lwd = 2)
plot(t, y, type = "l", lty = 3, lwd = 2)
plot(t, y, type = "l", lty = 4, lwd = 2)
plot(t, y, type = "l", lty = 5, lwd = 2)
```

### plot(..., bty = ...)
The box around the plot:
```{r, echo=TRUE}
bty = c("o", "n", "7", "L", "C", "U")
```
* o: complete box (default parameter)

* n: no box

* 7: top & right

* L: bottom & left

* C: top & left & bottom

* U: left & bottom & right
```{r, echo=TRUE}
t <- 1:20
y <- rnorm(20)
par(mfrow = c(2, 3)) ## split plot ranges
plot(t, y, type = "l", bty = "o", main = "bty = \" o \"")
plot(t, y, type = "l", bty = "n", main = "bty = \" n \"")
plot(t, y, type = "l", bty = "7", main = "bty = \" 7 \"")
plot(t, y, type = "l", bty = "L", main = "bty = \" L \"")
plot(t, y, type = "l", bty = "C", main = "bty = \" C \"")
plot(t, y, type = "l", bty = "U", main = "bty = \" U \"")
```

### plot(..., col = ...)
Use **col** for colors.

* Use number: **col = c(1, 2, 3, 4, 5, 6, 7, 8)**

* Use name: **col = c("red", "green", "blue", "grey")**

* Use hexadecimal code: **col = c("#009999", "#0000FF", ...)**

* More on colors: help(palette)
```{r, echo=TRUE}
t <- 1:20
y <- rnorm(20)

par(mfrow = c(2, 3)) ## split plot ranges

plot(t, y, type = "l", col = 1)
plot(t, y, type = "l", col = 2)
plot(t, y, type = "l", col = "green")
plot(t, y, type = "l", col = "grey")
plot(t, y, type = "l", col = "#009999")
plot(t, y, type = "l", col = "#0000FF")
```

### plot(..., main = ..., sub = ..., xlab = ..., ylab = ...)
Names of the plot:

* main: main title (at the top)

* sub: sub title (at the bottom)

* xlab: name of x-axis

* ylab: name of y-axis
```{r, echo=TRUE}
t <- 1:20
y <- rnorm(20)

par(mfrow = c(2, 3)) ## split plot ranges

plot(t, y, type = "l")
plot(t, y, type = "l", main = "Time series")
plot(t, y, type = "l", main = "Time series", sub = "this year")
plot(t, y, type = "l", main =  "Time series", xlab = "time")
plot(t, y, type = "l", main =  "Time series", xlab = "time", ylab = "")
plot(t, y, type = "l", main =  "Time series", xlab = "time", ylab = "return")
```

### plot(..., xlim = ..., ylim = ...)
Ranges of variables

* xlim = c(a, b): $a\leq x\leq b$

* ylim = c(a, b): $a\leq y\leq b$
```{r, echo=TRUE}
t <- 1:20
y <- rnorm(20)

par(mfrow = c(2, 2)) ## split plot ranges
plot(t, y, type = "b")
plot(t, y, type = "b", xlim = c(0, 50))
plot(t, y, type = "b", ylim = c(-5, 5))
plot(t, y, type = "b", xlim = c(0, 50), ylim = c(-5, 5))
```

### curve(...)
Instead using plot(x, y, type = "l"), use

<center>curve(expression, from = ..., to = ...)</center>

to draw a function of a variable.
```{r, echo=TRUE}
par(mfrow = c(2, 3))
curve(exp(-x), 0, 10)
curve(dnorm(x), -10, 10)
curve(x^2, -10, 10)
curve(sin, 0, 10)
curve(3*exp(-3*x), 0, 6)
curve(5*x+2, 0, 10)
```

## Low-level plotting functions: Add some objects to the existing plot
### Add points or lines to an existing plot

* **points(...)**: add points

* **lines(...)**: add lines

* **abline(...)**: add straight lines

<center>abline(a = ..., b = ..., h = ..., v = ..., reg = ..., coef = ...)</center>
```{r, echo=TRUE}
plot(t, y, type = "b")
abline(h = 0)
abline(v = 15, lty = 2, col = 8, lwd = 2)
points(t[15], y[15], pch = 19, col = 2)

x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)

plot(x, y, type = "p", pch = 19)
abline(a = 0, b = 3)

points(x, y1, col = 4, pch = 17)
abline(a = 0, b = 5, col = 4)
```

### Add text to an existing plot
Arbitrary locations:

* **text(x, y, labels = ...)**

Notations of objects in the plot:

* **legend(x, y, legend, ...)**
```{r, echo=TRUE}
x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)

plot(x, y, type = "p", pch = 19)
abline(a = 0, b = 3)

par(new=TRUE)
plot(x, y1, col = 4, pch = 20)
abline(a = 0, b = 5, col = 4)

text(-1.5, 2, "A", cex = 5)
text(1, -5, "B", cex = 5, col = 2)
legend("topleft", c("Y", "Y1"), col = c(1, 4), lty = c(1, 1), cex = 2)
```

### Re-arrange the region of a figure
Use **par(...)** to creat an arbitrary region

* mfrow = c(2, 2): Define a multi-figure display by row

* mfcol = c(2, 2): Define a multi-figure display by column

* mai, mar = c(3, 3, 1, 1): The four margin sizes, in inches (mai), or in text line units (mar).
```{r, echo=TRUE}
x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)

par(mfrow = c(1, 1), mar = c(0.5, 0.5, 0.5, 0.5))

plot(x, y, type = "p", pch = 19)
abline(a = 0, b = 3)

par(new=TRUE)
plot(x, y1, col = 4, pch = 20)
abline(a = 0, b = 5, col = 4)

text(-1.5, 2, "A", cex = 5)
text(1, -5, "B", cex = 5, col = 2)
legend("topleft", c("Y", "Y1"), col = c(1, 4), lty = c(1, 1), cex = 2)
```

### Split the whole plot by some locations of plot
Use **layout(...)**

* layout(matrix(c(1, 2, 3, 3), 2, 2, byrow = TRUE))
```{r, echo=TRUE}
x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)
y2 <- -2*x + rnorm(50, 0, 2)

layout(matrix(c(1, 2, 3, 3), 2, 2, byrow = TRUE))

plot(x, y, type = "p", pch = 19, main = "Y")
abline(a = 0, b = 3)

plot(x, y1, type = "p", pch = 19, main = "Y1")
abline(a = 0, b = 5)

plot(x, y2, type = "p", pch = 19, main = "Y2")
abline(a = 0, b = -2)
```

### Reference and packages
* http://web.ntpu.edu.tw/~cflin/Teach/R/R06EN06Graphics.pdf

* Package: ggplot2

* Package: ggmap (or some related to map)