---
title: "Related to Data"
author: "H24076150"
date: "2020/5/3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Frame
### data.frame()
* Most data sets are stored in R as data frames.

* These are like matrices, but with the columns having their own names.

* Columns can be of different types from each other.

Use the **data.frame()** to construct data frames from vectors:

Example:
```{r, echo=TRUE}
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
colors.and.numbers <- data.frame(colors, numbers, more.numbers = c(4, 5, 6))
colors.and.numbers

```
### Difference from matrix
* These are like matrices, but with the columns having their own names.

* Columns can be of different types from each other.

Use matrix:
```{r, echo=TRUE}
X <- matrix(NA, 3, 3)
X[, 1] <- c("red", "yellow", "blue")
X[, 2] <- c(1, 2, 3)
X[, 3] <- c(4, 5, 6)
colnames(X) <- c("colors", "numbers", "more.numbers")
X
```
### Type in a data fram
Use **str()** to examine the types of columns in the data frame.
```{r, echo=TRUE}
str(colors.and.numbers)
```
### Type in a data fram: character
Change the type from factor to character by using **stringAsFactors = FALSE**.
```{r, echo=TRUE}
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
colors.and.numbers <- data.frame(colors, numbers, more.numbers = c(4, 5, 6), stringsAsFactors = FALSE)
str(colors.and.numbers)
```
### More on data.frame(): add rows
Add column and row vectors into a data frame.

* row: use **rbind()**

* column: use the sign **$** or **cbind()**
```{r, echo=TRUE}
data1 <- data.frame(colors, numbers, more.numbers = c(4, 5, 6))
data1 <- rbind(data1, c("black", 3, 5))
data1
```
### stringsAsFactors = FALSE
```{r, echo=TRUE}
data2 <- data.frame(colors, numbers, more.numbers = c(4, 5, 6))
data2 <- rbind(data2, c("black", 3, 5))
data2
str(data2)
```
### Comparison of two types
```{r, echo=TRUE}
str(data1)
str(data2)
```
### More on data.frame(): add columns
Add column and row vectors into a data frame.

* column: use the sign **$** or **cbind()**

* row: use **rbind()**
```{r, echo=TRUE}
colors <- c("red", "yellow", "blue")
number1 <- c(1, 2, 3)
number2 <- c(10, 11, 12)
colors.and.numbers <- data.frame(colors, number1) ## construct a data frame
colors.and.numbers$more.numbers <- c(4, 5, 6) ## add one column
colors.and.numbers <- cbind(colors.and.numbers, number2) ## add one column
colors.and.numbers
```
## Data input and output at hard disk
### Change and find directory
Change the directory: **setwd("dir")**

Find the directory: **getwd()**

### Save and load the objects
Suppose you have constructed some R objects, type **dump(objects, file = "XXX.R")**.
```{r, echo=FALSE}
# Clear environment
rm(list = ls())
```

```{r, echo=TRUE}
X <- seq(0, 10, 0.1)
A <- 3
B <- mean(X)
dump(list = objects(), file = "all.R")
dump(c("A", "X"), file = "parts.R")
```
To retrieve the objects in a fucture session, type **source("XXX.R")**
```{r, echo=TRUE}
source("all.R")
source("parts.R")
```
* To save all objects in the current workspace, type **save.image("XXX.RData")**.

* To save parts of objects, type **save(a, b, c, file = "XXX.RData")**.

* Use **load("XXX.RData")** to retrieve data.
```{r, echo=FALSE}
# Clear environment
rm(list = ls())
```

```{r, echo=TRUE}
X <- seq(0, 10, 0.1)
A <- 3
B <- mean(X)
save.image("all.RData")
save(A, B, file = "parts.RData")
load("parts.RData")
load("all.RData")
```
### Save data to a file
You can save a matrix, a vector, or a data frame into two types of file, .txt(.dat) and .csv.

* Use **write.table(A, file = "XXX.txt")**

* Use **write.csv(A, file = "XXX.csv")**
```{r, echo=FALSE}
# Clear environment
rm(list = ls())
```
Save a data frame into a file:
```{r, echo=TRUE}
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
X <- data.frame(colors, numbers, more.numbers = c(4, 5, 6))
write.table(X, "matrix.txt")
write.csv(X, "matrix.csv")
```
### Read data from a file
You can load a matrix, a vector, or a data frame from two types of file, .txt(.dat) and .csv.

* Use **read.table(file = "XXX.txt")**

* Use **read.csv(file = "XXX.csv")**
```{r, echo=TRUE}
read.table("matrix.txt")
read.csv("matrix.csv")
```
If you read a data from a file with names of columns, you will get different input data of read.table() from that of read.csv().

With names of columns in a .txt file, you should add **header = TRUE**.

## R Package
### Introduction to R packages
* Other than some basic functions, some developed methodologies and tools are provided in a way of packsges. You can download them from CRAN or Github.

* To download a package, type **install.packages("XXX")**.

* Before using some functions in a downloaded package, you should call the package by typing **library(XXX)**.

Try package: data.table as follows.
```{}
install.packages("data.table")
```
### Use the functions in the installed packages
* Before using some functions in a downloaded package, you should call the package by typing **library(XXX)**.

Try the function **fread()**, which is similar to **read.table()** but faster and more convenient.
```{r, results="hide"}
library(data.table)
```

```{r, echo=TRUE}
mydat <- fread('test.csv')
head(mydat)
```
### Read data though internet
Data: http://www.stats.ox.ac.uk/pub/datasets/csb/ch11b.dat
```{r, echo=TRUE}
input <- fread("http://www.stats.ox.ac.uk/pub/datasets/csb/ch11b.dat")
head(input)
```
### More on packages
https://cran.r-project.org/web/packages/

https://www.r-bloggers.com/october-2019-top-40-new-r-packages/

## Tools for help
### Online help
When you know the name of a function and you want to see how to use it, type **help(name)** or **?name**.

For example:
```{r, echo=TRUE}
help(mean)
?median
```
### Built-in examples
A useful alternative to **help()** is the **example()** function.
```{r, echo=TRUE}
example(mean)
```
### Only knowing a keyword of function
For finding help when you don't know the function name, type **help.search()**.
```{r, echo=TRUE}
help.search("optimization")
```