---
title: "Vector"
author: "H24076150"
date: "2020/4/30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Numeric Vector
### Creating a vector
Creat two vectors and add one vector to another.
```{r, echo=TRUE}
vt1 <- c(1:5)
vt2 <- c(3:6)
vt2 <- c(vt2, vt1)
vt2
```
### Simple patterned vectors 
Use "seq()" and "rep()" to construct a vector with similar patterns.
```{r, echo=TRUE}
seq(1, 5, by = 0.5)
seq(1, 10, length.out = 5)
rep(3, 12)
rep(seq(2, 10, 2), 2)
```
### Creat a null or initial vector
```{r, echo=TRUE}
a <- c(3, 5, 9, 10)
b <- c(14, 15, 17)

v1 <- c()          ### 1st version
v1 <- c(v1, a, b)
v1

a <- c(3, 5, 9, 10)
b <- c(14, 15, 17)

v2 <- rep(0, 7)    ### 2nd version
v2[1:4] <- a
v2
v2[5:7] <- b
v2

a <- c(3, 5, 9, 10)
b <- c(14, 15, 17)

v3<- rep(NA, 7)   ### 3rd version
v3[1:4] <- a
v3
v3[5:7] <- b
v3
```
### Elements in the vector
Let b = {3,3,3,3,5,5,1,2,10,10,8}
```{r, echo=TRUE}
b <- c(3, 3, 3, 3, 5, 5, 1, 2, 10, 10, 8)
```
* Use "unique()" to return the elements that appear at least one time.
```{r, echo=TRUE}
unique(b)
```
* Use "table()" to count the number of each elememt from the vector.
```{r, echo=TRUE}
table(b)
```
### Set operations
```{r, echo=TRUE}
B <- c(3, 1, 5, 4, 7); C <- c(23, 1, 44, 5, 3)
union(B, C)
intersect(B, C)
setdiff(B, C)
setequal(B, C)
```
### Sorting and ordering of a vector
From small to large values, use "sort()" or "order()" to sort a vector or get the position of sorting a vector.

* sort(x, decreasing = FALSE)

* order(..., na.last = TRUE, decreasing = FALSE)
Let a = {3,4,2,7,10,23,8,9,11}

* Increasing: {2,3,4,7,8,9,10,11,23}

* Decreasing: {23,11,10,9,8,7,4,3,2}
```{r, echo=TRUE}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
sort(a)
sort(a, decreasing = TRUE)
```
### Get the position of sorting of a vector
Get the position (index) of sorting of a vector

* order(..., na.last = TRUE, decreasing = FALSE)
```{r, echo=TRUE}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
inc.order <- order(a)
dec.order <- order(a, decreasing = TRUE)
inc.order
a[inc.order]
dec.order
a[dec.order]
```
### Change values in a vector
Let a = {3,4,2,7,10,23,8,9,11}. Change the 5th element to -3 and the 8th element to -1.
```{r, echo=TRUE}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
a[5] <- -3
a[8] <- -1
```
### Deleting elements from a vector
We want to select all but the some elements of the vector as follows:
```{r, echo=TRUE}
vt1
vt1[-2]
vt2[-c(3:6)]
```
## Character vectors
### Print characters
```{r, echo=TRUE}
"Hello World!!!"
print("Hello R World")
```
### Split a string
```{r, echo=TRUE}
TEXT1 <- "Let's go to a game."
strsplit(TEXT1, " ")
TEXT2 <- "Hurry! We are late!"
strsplit(TEXT2, "!")
```
### Lowercase and uppercase
* toupper(): change to uppercase

* tolower(): change to lowercase
```{r, echo=TRUE}
toupper(TEXT1)
tolower(TEXT2)
```
### Character vectors
* Character vector: the vector is made up of strings of characters.

* All elements of a vector must be of the same type.
```{r, echo=TRUE}
colors <- c("red", "yellow", "BLUE")
more.colors <- c(colors, "green", "cyan")
more.colors
```
* An attempt to mix data types in a vector
```{r, echo=TRUE}
z <- c("red", "green", 12)
z
```
### Basic operations for character vectors
* To take substrings, use "substr()".
```{r, echo=TRUE}
colors <- c("red", "yellow", "blue")
substr(colors, 1, 3)
```
* To search a string from a vector, use "grep()", which is "Global Regular Expression Print". It returns the locations of the desired patterns.
```{r, echo=TRUE}
grep("y", more.colors)
grep("e", more.colors)
```
* How to show the whole character from grep()?
```{r, echo=TRUE}
y.in.V <- grep("y", more.colors)
e.in.V <- grep("e", more.colors)
more.colors[y.in.V]
more.colors[e.in.V]
```
* Is "grep()" case-sensitive?
```{r, echo=TRUE}
E.in.V <- grep("E", more.colors)
more.colors[E.in.V]
```
* To building up strings, use "paste()".
```{r, echo=TRUE}
paste(colors, "line")             ##with space
paste(colors, "line", sep = "")   ## no space
# use ";" to seperate strings
paste("I like",colors, collapse = "; ")
```
### Factors
factor(): factor(x = character())
```{r, echo=TRUE}
colors <- c("red", "yellow", "blue")
f.c <- factor(colors)
f.c
levels(f.c)
as.integer(f.c)
```
## Logical vector
### Logical words
TRUE and FALSE are reserved words denoting logical constants in the R, whereas T and F are global variables whose initial values set to these.
```{r, echo=TRUE}
TRUE
FALSE
T
F
```
### Logical number
Binary operators which allow the comparison of values in atomic vectors (relational operators).

* Number: Compare 3 and 5

* "==": equal

* "!=": not equal
```{r, echo=TRUE}
c(3 < 5, 3 > 5)
c(3 <= 5, 3 >= 5)
3 == 5
3 != 5
```
### Boolean operations
```{r, echo=TRUE}
A <- c(T, T, F, T, F); B <- c(3, 1, 5, 4, 7)
A
B[A]
!A
B[!A]
```
How many TRUE (or FALSE) in the vector?
```{r, echo=TRUE}
A <- c(T, T, F , T, F)
sum(A)
sum(!A)
```
Why can't we use the following code?
```{r, echo=TRUE}
length(A == TRUE)
```
How to modify it?
```{r, echo=TRUE}
length(A[A == F])
```
### Logical vector
Compare the same location of elements individually.
```{r, echo=TRUE}
B <- c(3, 1, 5, 4, 7); C <- c(23, 1, 44, 5, 3)
D <- c(2, 5, 6, 10)
B == D
```
### "is." series
"is.XYZ()" is any comparison to the type "XYZ" of an object.
* is.null()
* is.na()
* is.nan()
* is.infinite()
* is.character()
* is.integer()
* is.numeric()
```{r, echo=TRUE}
is.null(c())
c(is.na(NA), is.na(2))
x1 <- 1:5
x2 <- c(1:3, 2.2)
x3 <- colors
c(is.integer(x1), is.numeric(x1), is.double(x1))
c(is.integer(x2), is.numeric(x2), is.double(x2))
c(is.character(x3), is.numeric(x3))
```
### Examine the type of a vector
USe "class()" to identify the type of a vector.
```{r, echo=TRUE}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
b <- c("red", "yellow", "blue")
c <- c(TRUE, FALSE)
d <- f.c
c(class(a), class(b), class(c), class(d))
c(class(2), class(2.0), class(as.integer(2)))
```